Shape Flow - Deformable Models without Landmarks

### 2015-03-27 contents/implementation.tex
changes in pipeline description

We propose a novel framework for building deformable models that does not require any consistent set of annotated landmarks and is based on dense shape representation. Our method only requires a set of point or curve annotation that does not need to be consistent over different training samples.

It combines NICP, SVS representation and multi-image subspace flow in an effective framework that has significant descriptive power. The pipeline contains two major modules. The first module handles inconsistent annotation set by converting to landmark independent shape discriminator. While the other module produces shape flow on object discriminators to generate dense flow transformations in shape space following by robust PCA\cite{?} to generate deformable model. In this section, we present the entire architectures, design decision and algorithms.

\subsection{Sparse Landmarks Pre-Processing}
The first module is to bridge the gap that training data has to have consistent number of landmarks. Currently existing annotated databases contains large variety on landmarking, which largely restricted model training to stick to one data set also make evaluation challenging on using distinct data set. We propose an algorithm that utilizes a combination of Iterative Closest Point and Support Vector Shape for pre-processing before building the model.

\subsubsection{Model Invariance}
The model built should only captures deformations that specific to the object, transformations thus rotation, translation and scaling are processed to be invariant to the model\cite{?}. ICP is applied to align sparse landmarks. ICP is a perfect choice to align various point clouds with no restriction on shape dimension or quantity.

\subsubsection{Landmark Inconsistency}
Ordinarily, construction of deformable model requires training data set having consistent number of landmarks. But unavoidable changes has to be made before applying same algorithm on diverse annotated data set.

@@

\subsubsection{Trajectory Basis} \label{sec:trabasis}
Since shapes from training data set having completely no motion smoothness, it is of importance to contain constrains in optimization on how pixels moves from one shape frame to another. We constrain the optimization on trajectory subspace.
We supervise trajectories from sparse landmarks. To begin with, Non-rigid Iterative Closest Point (NICP)\cite{?} applied on sparse annotated points. NICP deforms target points to fit the template iteratively until all points on target deformed to match points on template. Performs the algorithm for all training data will get an implicit point correlation between sparse shapes.

@@

\subsubsection{Algorithm}
To register all decision functions to template shape, decision function $d_i(\bm{x}), i \in {1,...,F}$ are grouped into one sequence before applying flow algorithm. The objective cost function we would like to minimise is:
\begin{align}
    \operatorname*{arg\,min}_{\bm{u_n}(\bm{x}), \bm{v}}&=\alpha \int_{\Omega}\sum_{n=1}^F|\bm{d_n}(x+\bm{u_n}(x))-\bm{d_0}(x)\| dx  \label{eq:costfunc}\\
    &+ \beta \int_{\Omega}\sum_{n=1}^F\|\bm{u_n}(x)-\sum_{i=1}^R\bm{q_i}(n)\bm{v_i}(x)\|^2 dx \label{eq:lowrank}\\
    &+ \sum[\bm{TV}(Qv)]  + v^T.L.v
\end{align}
where $d_n(x)$ is the decision function from~\eqref{eq:decisionfunc}, which returns possibilities of given coordinate classified as shape component where coordianates are from set $\Omega \in \Re^2$. $TV(Qv)$ is total variation as regularization on low rank subspace, $Q$ is trajectory basis and $v^T.L.v$ is low rank spacial constrains.

>>>>

We propose a novel framework for building deformable models that does not require any consistent set of annotated landmarks and is based on dense shape representation. Our method only requires a set of point or curve annotation that does not need to be consistent over different training samples. It combines NICP, SVS representation and multi-image subspace flow in an effective framework that has significant descriptive power.

The proposed pipeline is depicted in Figure .... It takes as input a set of training images of an object class, along with point and/or curve annotations. Its first step is the application of ICP to achieve an initial alignment of the training images. Next, the shapes that correspond to the training data are consistently represented using multichannel SVS. These are then fed into a multi-image subspace flow estimation that establishes dense correspondences between all shapes of the training set. In order for the flow estimation to be accurate and exploit the correlation across the different training shapes, it utilises a correspondence basis that is built on sparse correspondences given by Nonrigid ICP alignment of the annotated data. Finally, the dense correspondences that are yielded by the optical flow serve as automatic dense landmarks and are used by a dense AAM training and fitting procedure. The following sections discuss in more detail the steps of the proposed pipeline.

%two major modules. The first module handles inconsistent annotation set by converting to landmark independent shape discriminator. While the other module produces shape flow on object discriminators to generate dense flow transformations in shape space following by robust PCA\cite{?} to generate deformable model. In this section, we present the entire architectures, design decision and algorithms.

\subsection{Shape Representation Without Consistent Annotations}
The first module is to bridge the gap that training data has to have consistent number of landmarks. Currently existing annotated databases contains large variety on landmarking, which largely restricted model training to stick to one data set also make evaluation challenging on using distinct data set. We propose an algorithm that utilizes a combination of Iterative Closest Point and Support Vector Shape for pre-processing before building the model.

\subsubsection{Initial Alignment of the Training Images}

The model built should only captures deformations that specific to the object, transformations thus rotation, translation and scaling are processed to be invariant to the model\cite{?}. ICP is applied to align sparse landmarks. ICP is a perfect choice to align various point clouds with no restriction on shape dimension or quantity.

\subsubsection{Multichannel Support Vector Shape Representation}
Ordinarily, construction of deformable model requires training data set having consistent number of landmarks. But unavoidable changes has to be made before applying same algorithm on diverse annotated data set.

@@

\subsubsection{Non-rigid ICP and Correspondence Basis Creation} \label{sec:trabasis}
Since shapes from training data set having completely no motion smoothness, it is of importance to contain constrains in optimization on how pixels moves from one shape frame to another. We constrain the optimization on correspondence subspace.
We supervise trajectories from sparse landmarks. To begin with, Non-rigid Iterative Closest Point (NICP)\cite{?} applied on sparse annotated points. NICP deforms target points to fit the template iteratively until all points on target deformed to match points on template. Performs the algorithm for all training data will get an implicit point correlation between sparse shapes.

@@

\subsubsection{Multi-image Subspace Flow}
To register all decision functions to template shape, decision function $d_i(\bm{x}), i \in {1,...,F}$ are grouped into one sequence before applying flow algorithm. The objective cost function we would like to minimise is:
\begin{align}
    \operatorname*{arg\,min}_{\bm{u_n}(\bm{x}), \bm{v}}&=\alpha \int_{\Omega}\sum_{n=1}^F|\bm{d_n}(x+\bm{u_n}(x))-\bm{d_0}(x)\| dx  \label{eq:costfunc}\\
    &+ \beta \int_{\Omega}\sum_{n=1}^F\|\bm{u_n}(x)-\sum_{i=1}^R\bm{q_i}(n)\bm{v_i}(x)\|^2 dx \label{eq:lowrank}\\
    &+ \sum[\bm{TV}(Qv)]  + v^T.L.v
\end{align}
where $d_n(x)$ is the decision function from~\eqref{eq:decisionfunc}, which returns possibilities of given coordinate classified as shape component where coordianates are from set $\Omega \in \Re^2$. $TV(Qv)$ is total variation as regularization on low rank subspace, $Q$ is trajectory basis and $v^T.L.v$ is low rank spacial constrains.
